---
title: "DiurnalMRI Methods Walkthrough"
output: pdf_document
date: "`r Sys.Date()`"
bibliography: '`r here::here("references.bib")`'
---

# Overview

The following page includes the full "data analysis and statistical methods" text with inline examples of code which executes an example of each method/statistic. We start by loading the functions into R:

```{r}
# Set the environment with relevant packages and load custom functions
source("code/setup.R")
```

# Subject-level oscillations (S-cosinor)

From methods:

> For each subject, cosinor linear regression was fitted with a 24-hr period to detect diurnal oscillations [@Cornelissen2014-zw]. The data were modeled by linear cosine and sine transformations of MRI acquisition time (t) to arrive at a model for MESOR, amplitude, and acrophase, starting with the following formula:

$$
y = \beta_0 + \beta_1 \cos(\frac{2{\pi}t}{\tau}) +  \beta_2 \sin(\frac{2{\pi}t}{\tau}) + error
$$

> where t is MRI acquisition time, $\tau$ is the period (24-hr). The intercept coefficient ($\beta_0$) is the MESOR. $\beta_1$ and $\beta_2$ are regression coefficients that were used to calculate the amplitude (A) and acrophase as:

$$
A = \sqrt{\beta_1^2 + \beta_2^2}
acrophase = atan2(\beta_1, \beta_2) \frac{\tau}{2\pi}
$$

> Such that $y = \beta0 + A \cos((t - acrophase)2\pi/\tau)$ + error. Each subject's p-values for the significance of diurnal oscillations were determined by the F-test comparing the cosinor model to an intercept-only (null) model and were used for the acrophase-agnostic test (described below). R2 of these within-subject models were obtained for each subject and used for the CBF differential oscillation strength test.

We use R package `DiscoRhythm`[@Carlucci2019] to obtain these cosinor statistics with `discoODAs`. Below is a minimal example of cosinor analysis on a simulated nine time-point dataset from a single subject.

```{r}
# starting at 8am for 24hrs
t = seq(8,32,3)
# setting parameters
b0=1 
acrophase=18 
A=1
# Observations
set.seed(1) # reproducibility
y = b0 + A*cos((t-acrophase)*2*pi/24) + rnorm(length(t))

# Here we format into SummarizedExperiment form, despite testing only a single feature 
# (i.e. subject/region/metric) as it will come in useful later for multi-feature 
# testing
suppressPackageStartupMessages(library(DiscoRhythm))
suppressPackageStartupMessages(library(SummarizedExperiment))
x <- matrix(y,nrow=1) 
se <- SummarizedExperiment(assay = x,
                           colData = DataFrame(Time=t)) 
# Applying cosinor
discoODAs(se,method="CS")$CS
```

# Group-level oscillations (G-cosinor)

From methods:

> The population-mean cosinor [@Cornelissen2014-zw] approach was applied with the aim to make inferences regarding a populational average rhythm for each group, and is referred to as G-cosinor throughout the text. This is a two stage model where first stage estimates were obtained by fitting each subjects' data to the S-cosinor model described above. Next, across subjects, the coefficients were averaged to obtain: $\beta_0^\star$ (MESOR), $\beta_1^\star$ and $\beta_2^\star$, and the following group-level model:

$$
\hat{y}  = \hat{\beta}_0^\star + \hat{A}^\star cos(t - \hat{acrophase}^\star)
$$

> Where "\*" indicates a population mean estimate, and A\* and acrophase\* were derived from $\beta_1^\star$ and $\beta_2^\star$. In this procedure, S-cosinor [ $\beta_1$ , $\beta_2$ ] values were considered jointly. A single hypothesis test was performed with the null that the true population mean values are zero (i.e., [$\beta_1$ ,$\beta_2^\star$] = [0, 0]) or equivalently, that A=0. The p-value was obtained by the corresponding multivariate F-test [@Cornelissen2014-zw] and is referred to as G-cosinor p-value throughout the text. In the regional analyses, for each metric, correction for multiple comparisons was performed using the Benjamini-Hochberg false discovery rate (FDR) [@Benjamini1995-ms]. 95% confidence intervals for A\*, acrophase\*, and $\beta_0^\star$ were estimated by population-mean cosinor methods [@Bingham1982-dc] deriving from the multivariate F-test.

These methods are implemented within the `discoPMcos()` function implemented for this study. We tested for equivalence of results with those of prior implementations of this method (R package cosinor2; test documented in "Gcosinor_Implementation_Check.Rmd"), and `discoPMcos()` improves upon performance with vectorization (useful for performing many such tests). Source code shown below:

```{r}
discoPMcos
```

Below is an example of a single test performed on data obtained by repeating the simulation above for multiple subjects. The S-cosinor methods are applied to the 16 simulated subjects to obtain 16 S-cosinor results.

```{r}
# Repeating the above simulation to simulate multiple subjects
set.seed(1) # reproducibility
x <- do.call(rbind,lapply(1:16,\(i){
  t = seq(8,32,3)
  b0=1 
  acrophase=18 
  A=1
  y = b0 + A*cos((t-acrophase)*2*pi/24) + rnorm(length(t))
  x=data.frame(matrix(y,nrow=1))
  df <- data.frame(subject=i,x)
  colnames(df)[-1] <- t
  df
}))

se <- SummarizedExperiment(assay = as.matrix(x[,-1]),
                           colData = DataFrame(Time=t)) 
by_subject <- data.frame(subject = x[,1], discoODAs(se,method="CS")$CS)
by_subject
```

`discoPMcos` accepts these results as input to test for a significant group-level rhythm across subjects, returning all statistics described in the above methods section.

```{r}
pmr <- discoPMcos(by_subject)
pmr
```

## Variance explained

> Variance explained. A proportion of variance explained was calculated as R2UN=1-SSR/TSS where the G-cosinor model was used to obtain:

$$
SSR = (y-\hat{y})^2 
$$

$$
TSS = (y-\beta^\star_0)^2
$$

> R2DM was calculated identically to R2UN except y was first mean subtracted within each individual's data, effectively setting each subject's $\beta_0$ to zero and controlling for interindividual variation in $\beta_0$. Therefore $\beta_0^\star$ of the G-cosinor models was also set to zero and we proceeded to calculate R2DM. One oscillating ROI in the GM-MD combined analysis had a low magnitude negative R2UN estimate which was set to zero for clarity.

An additional custom `rsqGcos()` function joins the original data, and the G-cosinor results to obtain these statistics. Source code shown below:

```{r}
rsqGcos
```

Where `cosinor_pred` returns model predictions of the population-mean. Source code shown below:

```{r}
cosinor_pred
```

Formatting the data and the `discoPMcos` results for joining within `rsqGcos` (by roi and measure), returns the two statistics:

```{r}
# Converting matrix to long data.frame
orig_dat_as_df <- data.frame(subject= rep(1:16,9),
             value=unlist(x[,2:ncol(x)]),
             time=rep(t,each=16),
             roi="whole-brain",
             measure="simulation")

# Add columns to join by
pmr$measure="simulation"
pmr$roi="whole-brain"

rsqGcos(pmr,orig_dat_as_df) 
```

## Differential acrophase

> Differential acrophase. When BPD and control groups each had significant G-cosinor oscillations, a difference in their acrophases was tested with permutation testing. Acrophase\* was obtained for both BPD (acrophase\*BPD) and control (acrophase\*CONTROL) groups. The minimum difference between them (minimum circular arc length) was then calculated.

The short custom function `circdist()` finds the acrophase difference. Source code shown below:

```{r}
circdist
```

For example, the difference between a 24-hr acrophase of 1 and 23 is 2 (crossing over the zero line):

```{r}
circdist(1,23)
```

> A permuted null distribution was generated by obtaining the same acrophase difference 10,000 times after shuffling BPD/control labels among subjects and recalculating group-wise acrophase\* estimates. The two-sided permutation p-value was the proportion of instances where an absolute acrophase difference from a permuted trial was larger than the real difference.

This permutation analysis was implemented in the WB_BPD_diffs.Rmd notebook.

# Acrophase-agnostic test for asynchronous subject-level diurnal oscillations.

> We implemented a method to complement the G-cosinor test to better utilize the variance explained by each individual within-subject S-cosinor fit. The S-cosinor F-test provided significance for each subject's fitted cosinor curve. As such, individual tests were not influenced by the amplitude/acrophase of other subjects, and were evaluated only by the degree to which they explained the within-subject data. We considered the global null hypothesis that no subjects showed oscillations. Under this null, p-values should follow a uniform distribution, and Fisher's combined probability test [@Fisher1925-zx] considers a likely distribution of p-values (as they would be if the global null were true) with the statistic:

$$
\chi^2_{2k} \sim -2 \sum_{i=1}^k \log(p_i) 
$$

> Where it follows a $\chi^2$ distribution with 2k (k=number of subjects) degrees of freedom. The test is conditional on the independence of subjects and is unweighted with respect to within-subject sample size.

With the same S-cosinor results:

```{r}
by_subject
```

We apply fishers method using the `metap` package [@Dewey2022] `sumlog` to each biological feature, across subjects.

```{r}
metap::sumlog
```

```{r}
metap::sumlog(by_subject$pvalue)
```

# Differential acrophase variance

> Acrophase variance Utilizing standard circular statistics [@Rao_Jammalamadaka2001-dg], with a list of acrophases ($\theta$) in radians), an estimate of their mean and variance was obtained by representing each of them as a point, u(C,S), on the unit circle as:

$$
u(C,S) = (\cos(\theta),\sin(\theta)) 
$$

> The mean (centroid) u of these points was obtained as:

$$
\bar{u}(\bar{C},\bar{S}) = (\frac{1}{k} \sum_{i=1}^k \cos(\theta_i),\frac{1}{k} \sum_{i=1}^k \sin(\theta_i))
$$

> Then, the length of u is a measure of acrophase consistency:

$$
\bar{R} = \sqrt{\bar{C}^2 + \bar{S}^2}
$$

> where 1 indicates all acrophases were identical, and conversely, circular variance is 1-R and is bounded from 0-1. The mean acrophase was obtained from the direction of u as =atan2(C,S).

In R, we utilized complex number notation to calculate all related directional statistics:

```{r}
acrophase_variance = 1 - (acro2complex(c(12,13,14)) |> mean() |> abs())
```

Where `acro2complex` is a simple custom function to convert acrophase into complex form. Source code shown below:

```{r}
acro2complex
```

For example, random uniform acrophases produce an estimated acrophase variance near one.

```{r}
1 - (acro2complex(runif(100)*24) |> mean() |> abs())
```

> Differential acrophase variance. Acrophase variance for both BPD (1-RBPD) and control (1-RCONTROL) groups were calculated as above. The difference between them (BPD - controls) was calculated. A permuted null distribution was generated by obtaining the same acrophase variance difference 10,000 times after shuffling BPD and control labels among subjects and recalculating group-wise acrophase variance estimates. The one-sided permuted p-value was the proportion of instances where the permuted acrophase variance difference was greater than the real difference.

This permutation analysis was implemented in the WB_BPD_diffs.Rmd notebook.

# References and page metadata
