---
title: Figure 2
---

[Skip to the final figure](#figure-2)

Currently in post only:
Bottom legend
Equals in panel titles (cannot use with facet_wrap label_parsed)

# Data Preparation

Importing the whole brain cosinor statistics and merging into one table.

```{r setup}
# Child usage must import this data and libraries
source("code/setup.R")

meandf <- read_csv("data/available_upon_request/whole_brain_data.csv") |> filter(dx == 0)
WB_Scosinor <- read_rds("output/WB_S-cosinor.rds")
WB_Gcosinor <- read_rds("output/WB_G-cosinor.rds")
WB_AA <- read_rds("output/WB_AA.rds")

# Body weight results
weight_Scosinor <- read_rds("output/body_weight/weight_S-cosinor_dx0.rds")
weight_Gcosinor <- read_rds("output/body_weight/weight_G-cosinor_dx0.rds")
weight_AA <- read_rds("output/body_weight/weight_AA_dx0.rds")
weights <-
  read_csv("data/available_upon_request/weights_kg.csv") |>
  filter(dx == 0) |>
  mutate(ltime = ifelse(session_is_day2(session, time), time + 24, time))
```

# Figure 2a: Raw data

```{r fig.width=4,fig.height=4}
meandf <-
  meandf |>
  mutate(measure = factor(measure, levels = fig1_modord)) |>
  arrange(measure)

wb_raw_data_mainfig_gg <- function(meandf,
                                   WB_Scosinor,
                                   fig1_modord) {
  labmap <-
    meandf |> select(measure) |> unique() |> 
    ungroup() |>
    arrange(match(measure, fig1_modord)) |>
    mutate(label = factor(measure,
      ordered = T, levels = fig1_modord,
      labels = int2extlab[measure]
    )) |>
    select(measure, label) |>
    unique()


  meandf |>
    inner_join(WB_Scosinor) |>
    mutate(fitted = cosinor_pred(ltime, amplitude, acrophase, mesor)) |>
    inner_join(labmap) |>
    ggplot(aes(ltime, value, group = subject)) +
    guides(linetype = guide_legend(expression(p[S]))) +
    night_rect() +
    geom_segment(aes(xend = ltime, yend = fitted), color = "grey") +
    geom_point(size = 0.2, color = colors$raw_data) +
    geom_smooth(
      linewidth = 0.4, method = "lm", formula = y ~ sinpi(x / 12) + cospi(x / 12),
      color = colors$ssc, se = F
    ) +
    facet_grid(label ~ subject,
      scales = "free",
      labeller = label_parsed
    ) +
    theme(strip.text.y = element_text(angle = -90)) +
    theme(panel.spacing.x = unit(0, "lines")) +
    theme(panel.spacing.y = unit(0, "lines")) +
    theme(axis.text.x = element_text(size = 6)) +
    theme(
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(color = "grey60", size = 0.3),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_line(color = "black", size = 0.5),
      panel.border = element_rect(color = NA),
      plot.title = element_text(size = 6, hjust = 0.5)
    ) +
    ggtitle("subject") +
    xlab("Time of Day (hr)") +
    geom_vline(xintercept = 8, color = "grey90") +
    scale_y_continuous(
      breaks = function(x) c(x[1] + diff(x) * 0.3, x[2] - diff(x) * 0.3),
      labels = my_val_formatter,
      minor_breaks = function(x) x
    ) +
    scale_x_continuous(
      breaks = c(8, 16, 24),
      labels = c(8, "", 24),
      limits = c(8, 36)
    ) +
    coord_cartesian(xlim = c(8, 32)) +
    theme_condense()
}
f2a <- wb_raw_data_mainfig_gg(meandf, WB_Scosinor, fig1_modord)
f2a
```

# Figure 2b: Polar plot

Distribution of acrophase estimates from 16 individuals with acrophases estimated on n=9 scans. G-cosinor result overlaid with 95% CI.

```{r mri-polar-p, fig.width=4.5,fig.height=4}
# Impure function (fig1_modord)
wb_polar_mainfig_gg <- function(WB_Scosinor, WB_AA, WB_Gcosinor, fig1_modord) {
  labmap <-
    WB_AA |>
    inner_join(WB_Gcosinor |> ungroup()) |>
    ungroup() |>
    arrange(match(measure, fig1_modord)) |>
    mutate(label = factor(measure,
      ordered = T, levels = fig1_modord[fig1_modord %in% measure],
      labels = paste0(
        int2extlab[measure],
        ifelse(p < 0.05, "^`*`", ""), " - p[G] -",
        my_val_formatter(pvalue)
      )
    )) |>
    select(measure, label) |>
    unique()

  # For CI plotting
  # Expand CI into two segments when crossing 24:00
  thispcr <- WB_Gcosinor
  if (WB_Gcosinor$acro_u[1] > 24) {
    if (WB_Gcosinorpcr$acrophase > 24) thispcr$acrophase <- WB_Gcosinor$acrophase %% 24
    tmp <- rbind(thispcr, thispcr)
    tmp$acro_u[1] <- 24
    tmp$acro_l[2] <- 0
    tmp$acro_u[2] <- tmp$acro_u[2] %% 24
    thispcr <- tmp
  }
  if (WB_Gcosinor$acro_l[1] < 0) {
    if (WB_Gcosinor$acrophase > 24) thispcr$acrophase <- WB_Gcosinor$acrophase %% 24
    tmp <- rbind(thispcr, thispcr)
    tmp$acro_l[1] <- tmp$acro_l[1] %% 24
    tmp$acro_u[1] <- 24
    tmp$acro_l[2] <- 0
    tmp$acro_u[2] <- tmp$acro_u[2] %% 24
    thispcr <- tmp
  }

  WB_Scosinor |>
    inner_join(labmap) |>
    ggplot(aes(acrophase, 1 - pvalue)) +
    # Build plot area
    night_rect_early() +
    annotate(geom = "rect", xmin = 0, xmax = 24, ymin = 0, ymax = 1, alpha = 1, color = NA, fill = "whitesmoke") +
    annotate(geom = "rect", xmin = 0, xmax = 24, ymin = -Inf, ymax = 0, fill = "white") +
    annotate(
      geom = "segment", x = seq(0, 24, 3), y = 1.8, yend = 0, xend = seq(0, 24, 3),
      color = "grey", size = 0.3
    ) +
    scale_y_continuous(breaks = c(0, 1), limits = c(-0.5, 1.8)) +
    geom_hline(yintercept = 1, color = "black", size = 0.2) +
    geom_hline(yintercept = 0.5, color = "black", size = 0.2, linetype = 3) +
    geom_segment(aes(xend = acrophase, yend = 0), color = colors$ssc, alpha = 0.3) +
    # Colors$scc not accurate for some reason...
    geom_point(
      shape = 21, stroke = 0, fill = "darkorange4", alpha = 0.8,
      color = "transparent", size = 1.2
    ) +
    # G-cosinor
    geom_point(
      data = thispcr |> filter(pvalue < 0.05) |> inner_join(labmap),
      aes(acrophase, 1.4, color = ifelse(pvalue < 0.05, "p[G]<0.05", "p[G]>0.05")), size = 1.2
    ) +
    geom_errorbar(
      data = thispcr |> filter(pvalue < 0.05) |> inner_join(labmap),
      aes(
        xmin = acro_u, xmax = acro_l, y = 1.4,
        color = ifelse(pvalue < 0.05, "p[G]<0.05", "p[G]>0.05")
      ), width = 0.5
    ) +
    facet_wrap(label ~ ., nrow = 3, labeller = label_parsed) +
    theme(
      panel.border = element_blank(),
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.spacing.x = unit(0, "line")
    ) +
    coord_polar() +
    theme(
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_blank()
    ) +
    xlab("Acrophase (hr)") +
    ylab(expression(1 - p[S])) +
    scale_color_manual(
      labels = c(
        bquote(p[G] * ">0.05"),
        bquote(p[G] * "<0.05")
      ),
      values = c("p[G]>0.05" = "white", "p[G]<0.05" = colors$gcos_sig)
    ) +
    guides(color = guide_legend("")) +
    scale_x_continuous(limits = c(0, 24), breaks = seq(0, 24, 6)) +
    guides(size = guide_legend(expression(-log10(p[S])))) +
    theme(panel.spacing = unit(0, "lines")) +
    theme(
      panel.border = element_blank(),
      panel.background = element_blank(),
      # Legend
      legend.title.align = 0,
      legend.position = "right",
      legend.title = element_text(size = 8, vjust = 0, hjust = -3),
      panel.spacing.x = unit(1, "line"),
      panel.grid = element_blank()
    ) +
    theme_condense()
}

# (Partial) Source Data
fig2bSD <- 
  WB_Gcosinor |> 
  select(measure,pvalue,acrophase,acro_l,acro_u)
write_csv(fig2bSD |> mutate(measure=int2extlab[measure]),
          "output/source_data/Figure_2b_partial.csv")

f2b <- wb_polar_mainfig_gg(WB_Scosinor, WB_AA, fig2bSD, fig1_modord)
f2b
```

# Figure 2c: Unnormalized G-cosinor overlay

An additional illustration of the amplitude compared to interindividual variation

```{r fig.height=4,fig.width=4}
# library(ggpubr) TODO: what's it for?
wb_unnorm_gcos <- function(WB_Gcosinor, meandf, fig1_modord) {
  # Create label for each metric
  labmap <-
    WB_Gcosinor |>
    ungroup() |>
    arrange(match(measure, fig1_modord)) |>
    mutate(label = factor(measure,
      ordered = T, levels = fig1_modord,
      labels = paste0(
        int2extlab[measure], "-R[UN]^2-",
        my_val_formatter(rsq, doparse = FALSE)
      )
    )) |>
    select(measure, label) |>
    unique()

  # Getting data for gcos predictions
  gcossim <- WB_Gcosinor |>
    select(measure, MESOR, amplitude, acrophase, pvalue) |>
    group_by(measure) |>
    nest() |>
    mutate(tmp = map(data, function(x) {
      t <- 8:34
      data.frame(
        value = cosinor_pred(
          t, as.numeric(x[1, 2]),
          as.numeric(x[1, 3]), as.numeric(x[1, 1])
        ),
        ltime = t, subject = "Cohort", pvalue = x$pvalue
      )
    })) |>
    unnest(tmp)

  meandf |>
    inner_join(labmap) |>
    ggplot() +
    night_rect() +
    geom_point(aes(ltime, value), size = 0.4, color = colors$raw_data) +
    stat_smooth_cosinor(aes(ltime, value, group = subject), se = F, alpha = 0.5, color = colors$ssc) +
    geom_rect(
      data = WB_Gcosinor |> filter(pvalue < 0.05) |> inner_join(labmap),
      aes(
        xmin = 8, xmax = 39, x = 0, y = MESOR,
        ymin = MESOR - amplitude, ymax = MESOR + amplitude
      ),
      fill = "lightblue", alpha = 0.7
    ) +
    geom_errorbar(
      data = WB_Gcosinor |> filter(pvalue < 0.05) |> inner_join(labmap),
      aes(x = 36, y = MESOR, ymin = MESOR - amplitude, ymax = MESOR + amplitude),
      color = "blue", width = 1.75
    ) +
    geom_text(
      data = WB_Gcosinor |> filter(pvalue < 0.05) |> inner_join(labmap),
      aes(label = "2A", x = 36, y = MESOR + amplitude * 5), size = 2,
      color = "blue"
    ) +
    geom_line(
      data = gcossim |> inner_join(labmap),
      aes(ltime, value, color = ifelse(pvalue < 0.05, "p[G]<0.05", "p[G]>0.05")), size = 0.8
    ) +
    scale_color_manual(
      labels = c(
        bquote(p[G] * ">0.05"),
        bquote(p[G] * "<0.05")
      ),
      values = c("p[G]>0.05" = "black", "p[G]<0.05" = colors$gcos_sig)
    ) +
    guides(color = "none") +
    facet_wrap(label ~ ., scales = "free", labeller = label_parsed) +
    theme(panel.spacing = unit(0.1, "line")) +
    theme(panel.grid.minor = element_blank()) +
    xlab("Time of Day (hr)") +
    scale_x_continuous(
      breaks = c(8, 16, 24, 32),
      labels = c(8, "", 24, ""),
      limits = c(8, NA)
    ) +
    scale_y_continuous(labels = my_val_formatter) +
    coord_cartesian(xlim = c(8, 36)) +
    theme_condense()
}

# (Partial) Source Data
fig2cSD <- 
  WB_Gcosinor |> 
  select(measure,pvalue,amplitude,acrophase,MESOR,rsq)
write_csv(fig2cSD |> mutate(measure=int2extlab[measure]),
          "output/source_data/Figure_2c_partial.csv")

ampoverlay_p <- wb_unnorm_gcos(fig2cSD, meandf, fig1_modord)
ampoverlay_p
```

# Figure 2d: De-meaned G-cosinor {.tabset}

```{r fig.height=4,fig.width=4}
wb_norm_gcos <- function(WB_Gcosinor, meandf, fig1_modord) {
  labmap <-
    WB_Gcosinor |>
    ungroup() |>
    arrange(match(measure, fig1_modord)) |>
    mutate(label = factor(measure,
      ordered = T, levels = fig1_modord,
      labels = paste0(
        int2extlab[measure], " - R[DM]^2 :",
        my_val_formatter(rsqdm)
      )
    )) |>
    select(measure, label) |>
    unique()

  gcos_preds <- WB_Gcosinor |>
    mutate(fitted = list(data.frame(
      t = 8:34,
      y_pred = cosinor_pred(8:34, amplitude, acrophase)
    ))) |>
    unnest(fitted) |>
    inner_join(labmap)

  meandf |>
    group_by(measure, subject) |>
    mutate(value = scale(value, scale = FALSE)) |>
    inner_join(labmap) |>
    ggplot(aes(ltime, value, group = NA)) +
    night_rect() +
    geom_point(size = 0.3, color = colors$raw_data) +
    xlab("Time of Day (hr)") +
    ylab("Demeaned values") +
    scale_x_continuous(
      breaks = c(8, 16, 24, 32),
      labels = c(8, "", 24, ""),
      limits = c(8, 34)
    ) +
    scale_y_continuous(labels = my_val_formatter) +
    stat_smooth_cosinor(aes(group = subject), se = F, alpha = 0.5, color = colors$ssc) +
    geom_line(data = gcos_preds, aes(t, y_pred,
      color = ifelse(pvalue < 0.05, "p[G]<0.05", "p[G]>0.05")
    ), size = 1) +
    scale_color_manual(
      labels = c(
        bquote(p[G] * ">0.05"),
        bquote(p[G] * "<0.05")
      ),
      values = c("p[G]>0.05" = "black", "p[G]<0.05" = colors$gcos_sig)
    ) +
    guides(color = "none") +
    theme(panel.spacing = unit(0, "line")) +
    facet_wrap(nrow = 4, label ~ ., scales = "free", labeller = label_parsed) +
    theme_condense()
}

# (Partial) Source Data
fig2dSD <- 
  WB_Gcosinor |> 
  select(measure,pvalue,amplitude,acrophase,MESOR,rsqdm)
write_csv(fig2dSD |> mutate(measure=int2extlab[measure]),
          "output/source_data/Figure_2d_partial.csv")

Gcos_ondata_p <- wb_norm_gcos(fig2dSD, meandf, fig1_modord)
Gcos_ondata_p
```

# Weights

Reuse MRI metrics/code for weight

## Raw data

```{r fig.height=1.6, fig.width=4} 
rawlin_p <- wb_raw_data_mainfig_gg(
  weights |> mutate(measure = "weight", value = weight),
  weight_Scosinor, "weight"
)
rawlin_p
```

## Polar Acrophases

```{r fig.width=4,fig.height=1.4}
polar_weights_p <- wb_polar_mainfig_gg(
  weight_Scosinor |> mutate(measure = "weight"),
  weight_AA,
  weight_Gcosinor |> mutate(measure = "weight"),
  "weight"
)
polar_weights_p
```

## Un-normalized S/G cosinor fits

```{r fig.height=1.4,fig.width=1.4}
Gcos_un_weights_p <- wb_unnorm_gcos(
  weight_Gcosinor |> mutate(measure = "weight"),
  weights |> mutate(measure = "weight", value = weight),
  "weight"
)
Gcos_un_weights_p
```

## De-meaned S/G cosinor 

```{r fig.width=1.4,fig.height=1.4}
Gcos_weights_p <- wb_norm_gcos(
  weight_Gcosinor |> mutate(measure = "weight"),
  weights |> mutate(measure = "weight", value = weight),
  "weight"
)
Gcos_weights_p
```

# Figures {.tabset}

## Figure 2 {#figure-2}

```{r figure_2,fig.height=10,fig.width=8.3, dev=c("png","pdf")}
f2a +
  #-next
  f2b + guides(color = "none") +
  #-next
  ampoverlay_p + guides(color = "none") +
  #-next
  Gcos_ondata_p + guides(color = "none") +
  #-patchwork
  plot_layout(design = "AAABBB
				          AAABBB
				          AAABBB
				          AAABBB
				          AAABBB
				          CCCDDD
				          CCCDDD
				          CCCDDD
				          CCCDDD
				          ") +
  plot_annotation(tag_levels = "a") +
  plot_layout(guides = "collect") & theme_condense()
```

## Figure Sbody-weight

```{r fig.height=3}
  #-next
  rawlin_p + ylab("Body weight (kg)") +
  #-next
  polar_weights_p + guides(color = "none") +
  #-next
  Gcos_un_weights_p + ylab("Body weight (kg)") +
  #-next
  Gcos_weights_p+
  plot_layout(design = "AAAABCD") +
  plot_annotation(tag_levels = "a") +
  plot_layout(guides = "collect") & theme_condense()
```

