---
title: Figure 3 and Figure S3
---

```{r setup}
source("code/setup.R")

# Data prep
WB_Gcosinor <- read_rds("output/WB_G-cosinor.rds")
WB_Scosinor <- read_rds("output/WB_S-cosinor.rds")
# Gather results
ROI_Gcosinor <- readRDS("output/ROI_G-cosinor.rds")
ROI_Scosinor <- readRDS("output/ROI_S-cosinor.rds") |> unnest(bsc)

plotdf <-
  ROI_Scosinor |>
  # Calc summary stats
  group_by(measure,subject) |> 
  mutate(sub_resultant=acro2complex(acrophase) |> mean() |> abs()) |> 
  inner_join(WB_Scosinor |> select(measure,subject,sub_macro=acrophase)) |> 
  group_by(measure,roi) |> 
  rename(p_sub_cos = pvalue) |> 
  group_by(measure) |> 
  mutate(p_rank=-rank(p_sub_cos)/length(p_sub_cos)) |> 
  # Join to rest
	group_by(measure,roi) |> 
	nest(by_visit = c(-measure,-roi)) |>
	inner_join(ROI_Gcosinor |> 
			   	select(measure, roi, 
			   	       p_gcos_cos=pvalue,
			   	       gcos_acro = acrophase, acro_u, acro_l,
			   	       gcos_amp = amplitude,
			   	       q_gcos_cos = qvalue,
			   	       amp_u,
			   	       amp_l,
			   	       gcos_mesor = MESOR))

# Ordering ROIs and adjusting acrophases
tmpdf <-
	plotdf |>
	unnest(by_visit) |>
	ungroup() |> 
	group_by(measure) |> 
	nest() |> 
	mutate(tmp=map(data,function(x){
		x |> 
			mutate(roif=factor(roi)) |>
			mutate(subject=factor(subject)) |> 
			ungroup() |> 
			arrange(p_gcos_cos,(sub_macro)%%24) |> 
			mutate(roif=fct_inorder(roif)) |> 
			mutate(subject=fct_inorder(subject)) |> 
			mutate(subjectn=as.numeric(subject)) |> 
			mutate(roin=as.numeric(roif)) |> 
			mutate(roif2=fct_inorder(roif)) |> 
			mutate(roin2=as.numeric(roif2))
	})) |> unnest(tmp) |> 
	ungroup() |> 
	arrange(measure) |> 
	mutate(roin_comb=factor(paste(measure,roin))) |> 
	mutate(roin_comb=fct_inorder(roin_comb))
```

Applying some preprocessing to the data:

- Ordering ROIs and subjects
- Adjusting acrophases to a 9-33 timescale

# Figure 3a: Subject vs Acrophase {.tabset}

## Figure S3a: Tick plot

```{r f3a-sub-act, fig.width=3}
# https://github.com/dgrtwo/drlib/blob/master/R/reorder_within.R
# For applying unique orderings within each facet while retaining
# the original labels
reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}
scale_y_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_y_discrete(labels = function(x) gsub(reg, "", x), ...)
}
scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

binsize=0.1
cuts <- seq(0,24,binsize)
bins <- levels(cut(1,cuts))

f3a_sub_act <-
	tmpdf |>
  mutate(measure=recode(measure, !!!int2extlab)) |> 
  group_by(measure,subject,sub_macro) |> 
	summarize(acro_cnt = table(cut(acrophase,cuts)) |> enframe()) |> 
	unnest(acro_cnt) |> 
	mutate(tod=cuts[which(bins == name)]+binsize/2) |> 
	mutate(value=as.numeric(value)/sum(value)) |>
	filter(value!=0) |> 
	ggplot(aes(reorder_within(subject, (sub_macro)%%24, measure),tod))+
	night_recty_early()+
	night_recty()+
	scale_x_reordered()+
	xlab("Subject")+ylab("ROI S-cosinor Acrophase")+
  geom_tile(fill="black",width=0.5,height=binsize)+
	coord_flip()+
	facet_grid(measure~.,scales="free")+ #,space = "free")
  geom_point(data=WB_Scosinor |> 
               mutate(measure=recode(measure, !!!int2extlab)),
             aes(reorder_within(subject,(acrophase)%%24, measure),
                 acrophase,group=NA),size=1,alpha=0.7,color=colors$ssc,
             position=position_nudge(x = 0.3, y = 0))+
	scale_y_continuous(limits=c(0,24),breaks=seq(0,24,3))+
  theme_condense()
f3a_sub_act
```

## Figure 3a: Subset for main fig

```{r fig.width=2.8,fig.height=8.5}
binsize=0.1
cuts <- seq(0,24,binsize)
bins <- levels(cut(1,cuts))

# (Partial) Source Data
write_csv(WB_Gcosinor |> 
  filter(measure %in% c("cbf","gm_md","md")) |> 
  mutate(measure=recode(measure, !!!int2extlab)) |> 
  select(measure,acrophase),"output/source_data/Figure_3a_partial.csv")

f3a_sub_act_ss <-
	tmpdf |> filter(measure %in% c("cbf","gm_md","md","qt1")) |>
  # Keep measure ordering
  mutate(measure=recode(measure, !!!int2extlab)) |> 
  group_by(measure,subject,sub_macro) |> 
	summarize(acro_cnt = table(cut(acrophase,cuts)) |> enframe()) |> 
	unnest(acro_cnt) |> 
	mutate(tod=cuts[which(bins == name)]+binsize/2) |> 
  # get acrobins with data
	mutate(value=as.numeric(value)/sum(value)) |>
	filter(value!=0) |> 
	ggplot(aes(reorder_within(subject, (sub_macro)%%24, measure),tod))+
	night_recty_early()+
  night_recty()+
	scale_x_reordered()+
	xlab("Subject")+ylab("ROI S-cosinor Acrophase")+
  #Data
  geom_point(data=WB_Scosinor |> 
               filter(measure %in% c("cbf","gm_md","md","qt1")) |> 
               mutate(measure=recode(measure, !!!int2extlab)),
             aes(reorder_within(subject,(acrophase)%%24, measure),
                 acrophase,group=NA),size=1.5,alpha=0.7,color=colors$ssc,
             position=position_nudge(x = 0.3, y = 0))+
  geom_tile(fill="black",width=0.5,height=binsize)+
  coord_flip(ylim = c(0,24))+
	facet_grid(measure~.,scales="free")+ #,space = "free")
  geom_hline(data=WB_Gcosinor |> 
               filter(measure %in% c("cbf","gm_md","md")) |> 
               mutate(measure=recode(measure, !!!int2extlab)),
             aes(yintercept=acrophase,group=NA),
             size=1,color=colors$gcos_sig)+
	scale_y_continuous(breaks=seq(0,24,3))+
  theme_condense()+
  theme(axis.text=element_text(size=8))
f3a_sub_act_ss
```


# AcroVariance Summaries {.tabset}

## Figure S3b: rho within-subject summary

```{r fig.width=5}
rhosum_p <-
	plotdf |> 
  mutate(measure=recode(measure, !!!int2extlab)) |> 
	unnest(by_visit) |>
	ungroup() |> 
	select(measure,subject,sub_macro,sub_resultant) |> unique() |> 
	ggplot(aes(reorder_within(subject, (sub_macro)%%24, measure),
			   1-sub_resultant))+
	geom_col(width = 0.5,fill="#C26179")+
	coord_flip()+
	xlab(NULL)+
	ylab("Acrophase Variance")+
	scale_x_reordered()+
  scale_y_continuous(breaks=c(0,0.5,1))+
	facet_grid(measure~.,scales="free")+
	theme_condense()
rhosum_p
```

## Figure 3a.2: rho within-subject summary (subset)

```{r}
rhosum_ss_p <-
	plotdf |> 
	filter(measure %in% c("cbf","gm_md","md","qt1")) |>   
	unnest(by_visit) |>
	ungroup() |> 
	select(measure,subject,sub_macro,sub_resultant) |> unique() |> 
	ggplot(aes(reorder_within(subject, (sub_macro)%%24, measure),
			   1-sub_resultant))+
	geom_col(width = 0.5,fill="#C26179")+
	coord_flip()+
	xlab(NULL)+
	ylab(expression(1-abs(rho[within-subject])))+
	scale_x_reordered()+
  scale_y_continuous(breaks=c(0,0.5,1))+
	facet_grid(int2extlab[measure]~.,scales="free")+ #,space = "free")+
	theme_condense()
rhosum_ss_p
```


# Figure S3c: ROI vs Acro {.tabset}

For supplemental

```{r p-roiact, fig.width=4}
binsize=0.4
cuts <- seq(0,24,binsize)
bins <- levels(cut(1,cuts))

# (Partial) Source Data
figS3cSD <-
  tmpdf |>
  mutate(measure=recode(measure, !!!int2extlab)) |> 
  filter(q_gcos_cos<0.05) |>
  select(roin,roi,measure,tod=gcos_acro) |>
  mutate(subject="Cohort") |> unique()
write_csv(figS3cSD,"output/source_data/Figure_S3c_partial.csv")

p_roiact<-
  tmpdf |>
  group_by(measure,roin) |> 
  mutate(measure=recode(measure, !!!int2extlab)) |> 
	summarize(acro_cnt = table(cut(acrophase,cuts)) |> enframe()) |> 
	unnest(acro_cnt) |> 
	mutate(tod=cuts[which(bins == name)]+binsize/2) |> 
	mutate(value=as.numeric(value)/sum(value)) |> 
	filter(value!=0) |> 
	ggplot(aes(-roin,tod))+
	night_recty_early()+
	scale_x_reordered()+
	ylab("ROI S(G)-cosinor Acrophase")+
	xlab("ROI ranked by G-cosinor p")+
  geom_tile(fill="black",width=1,height=binsize)+
	coord_flip()+
	facet_grid(measure~.,scales="free")+ #,space = "free")
  geom_point(size=0.5,data=figS3cSD,color="blue",alpha=0.7)+
	scale_y_continuous(limits=c(0,24),breaks=seq(0,24,3))+
  theme_condense()
p_roiact
```

# Polar scatter G-cosinor {.tabset}

In main text, maximal acrophase differences are reported alongside the span of CIs

```{r}
plotdf |> 
  inner_join(WB_Gcosinor |> select(measure,acrophase),by="measure") |> 
  filter(q_gcos_cos<0.05) |> 
  group_by(measure) |> 
  summarize(circdist(gcos_acro,acrophase) |> abs() |> max())

WB_Gcosinor |>
  filter(pvalue<0.05) |> 
  mutate(span=circdist(acro_u,acro_l)) |> 
  select(measure,span)
```

## Figure 3c: Main

```{r gcrose-ss-p, fig.height=8, fig.width=2.75}
predat <- plotdf |>
	filter(measure %in% c("cbf","gm_md","md","qt1")) |> 
	group_by(measure) |> 
	arrange(measure) |> 
	mutate(nsig=sum(q_gcos_cos<0.05),n=n())

## Source Data
# ROI data
fig3cSD <- 
  predat |>
  ungroup() |>
  mutate(measure=int2extlab[measure]) |> 
  select(measure, nsig, n, p_gcos_cos,q_gcos_cos,gcos_acro)
write_csv(fig3cSD, "output/source_data/Figure_3c_part1.csv")
# WB data
fig3cSD2 <- 
  WB_Gcosinor |> 
  right_join(predat |> 
               select(measure,nsig,n) |> 
               unique()) |>
  filter(pvalue<0.05) |> 
  ungroup() |> 
  mutate(measure=int2extlab[measure]) |> 
  select(measure,nsig,n,pvalue,acrophase,acro_l,acro_u)
write_csv(fig3cSD2,"output/source_data/Figure_3c_part2.csv")


gcrose_ss_p <-
  fig3cSD |> 
  mutate(label=fct_inorder(paste0(measure," (",nsig,"/",n,")"))) |> 
	ggplot(aes(gcos_acro,-log10(p_gcos_cos)))+
	night_rect_early()+
	geom_rect(xmin=-1,xmax=25,ymin=-2,ymax=0,fill="white",color=NA)+
	annotate(geom="segment",x=seq(0,24,3),y=4,yend=0,xend=seq(0,24,3),
             color="grey90",size=0.3)+
  geom_hline(yintercept = 0,color="grey50")+
	geom_hline(yintercept = c(1,2,3,4),color="grey90",size=0.5)+
	coord_polar()+
	scale_y_continuous(limits = c(-2,NA),breaks=c(0,1,2,3,4))+
	scale_x_continuous(limits=c(0,24),breaks=seq(0,24,3))+
	xlab("ROI G-cosinor Acrophase")+
	ylab(expression(log[10](p[G])))+
	# Making WM-MD larger since fewer ROIs
  # unused
	geom_point(aes(size=1/n,fill=q_gcos_cos<0.05),alpha=0.7,
      shape=21, stroke=0,  
      color="transparent"
    )+
	scale_size_continuous(range = c(1,1),guide='none')+
  theme(panel.border=element_blank())+
	theme(panel.grid=element_blank())+
	geom_point(data=fig3cSD2 |> 
	             	mutate(label=fct_inorder(paste0(measure," (",nsig,"/",n,")"))),  
			   aes(acrophase,5,color=pvalue<0.05),size=2)+
	geom_errorbar(data=fig3cSD2 |> mutate(label=fct_inorder(paste0(measure," (",nsig,"/",n,")"))),  
			   aes(acrophase,5,xmin=acro_l,xmax=acro_u,
			   	color=pvalue<0.05),size=1.5)+
	scale_fill_manual(labels=c(bquote(q[G]*"<0.05"),
								bquote(q[G]*">0.05")),
								values = c("TRUE"=colors$gcos_sig,
								  "FALSE"="black"),guide="none")+
	scale_color_manual(labels=c(bquote(q[G]*"<0.05"),
								bquote(q[G]*">0.05")),
								values = c("TRUE"=colors$gcos_sig,
								  "FALSE"="black"),guide='none')+
  # Remove for post
  # guides(color=guide_legend(title=""))+
	facet_grid(label~.)+theme_condense()
gcrose_ss_p
```

## Figure S3d: Sro3 

```{r}
# Copied from main fig code and commented out
predat <- plotdf |>
	# filter(measure %in% c("cbf","gm_md","md","qt1")) |> 
	group_by(measure) |> 
	arrange(measure) |> 
	mutate(nsig=sum(q_gcos_cos<0.05),n=n())

## Source Data
# ROI data
figS3dSD <- 
  predat |>
  ungroup() |> 
  mutate(measure=int2extlab[measure]) |> 
  select(measure,nsig,n, p_gcos_cos,n, q_gcos_cos,gcos_acro)
write_csv(figS3dSD,"output/source_data/Figure_S3d_part1.csv")
# WB data
figS3dSD2 <- 
  WB_Gcosinor |> 
  right_join(predat |> 
               select(measure,n,nsig) |> 
               unique()) |>
  filter(pvalue<0.05) |>
  ungroup() |> 
  mutate(measure=int2extlab[measure]) |> 
  select(measure,nsig,n,acrophase,acro_l,acro_u,pvalue)
write_csv(figS3dSD2,"output/source_data/Figure_S3d_part2.csv")


gcrose_p <-
	figS3dSD |> 
  	mutate(label=fct_inorder(paste0(measure," (",nsig,"/",n,")"))) |> 
	ggplot(aes(gcos_acro,-log10(p_gcos_cos)))+
	night_rect_early()+
	geom_rect(xmin=-1,xmax=25,ymin=-2,ymax=0,fill="white",color=NA)+
	annotate(geom="segment",x=seq(0,24,3),y=4,yend=0,xend=seq(0,24,3),
             color="grey90",size=0.3)+
  geom_hline(yintercept = 0,color="grey50")+
	geom_hline(yintercept = c(1,2,3,4),color="grey90",size=0.5)+
	coord_polar()+
	scale_y_continuous(limits = c(-2,NA),breaks=c(0,1,2,3,4))+
	scale_x_continuous(limits=c(0,24),breaks=seq(0,24,3))+
	xlab("ROI G-cosinor Acrophase")+
	ylab(expression(log[10](p[G])))+
	# Making WM-MD larger since fewer ROIs
  # unused
	geom_point(aes(size=1/n,fill=q_gcos_cos<0.05),alpha=0.7,
      shape=21, stroke=0,  
      color="transparent"
    )+
	scale_size_continuous(range = c(1,1),guide='none')+
  theme(panel.border=element_blank())+
	theme(panel.grid=element_blank())+
	geom_point(data=figS3dSD2 |> 
	             	mutate(label=fct_inorder(paste0(measure," (",nsig,"/",n,")"))),  
			   aes(acrophase,5,color=pvalue<0.05),size=1.5)+
	geom_errorbar(data=figS3dSD2 |> 	mutate(label=fct_inorder(paste0(measure," (",nsig,"/",n,")"))),  
			   aes(acrophase,5,xmin=acro_l,xmax=acro_u,
			   	color=pvalue<0.05),size=1)+
	scale_fill_manual(labels=c(bquote(q[G]*"<0.05"),
								bquote(q[G]*">0.05")),
								values = c("TRUE"=colors$gcos_sig,
								  "FALSE"="black"),guide="none")+
	scale_color_manual(labels=c(bquote(q[G]*"<0.05"),
								bquote(q[G]*">0.05")),
								values = c("TRUE"=colors$gcos_sig,
								  "FALSE"="black"),guide='none')+
  guides(color=guide_legend(title=""))+
	facet_grid(label~.)+theme_condense()
gcrose_p
```


# Brain images {.tabset}

## Glasser

```{r}
library(ggsegGlasser)
library(ggseg)
meta <- get_glasser_spatial_meta()

meta[!(meta$region %in% glasser$data$region),]
glasser$data[!(glasser$data$region %in% meta$region),]
full_join(ROI_Gcosinor |> filter(measure=="cbf"),
          meta |> select(region,hemi,roi)) |>
  filter(!is.na(qvalue) & !is.na(region)) |> pull(roi)

# Source Data
fig3bSD <- 
  ROI_Gcosinor |>
  # GM with any(qvalue<0.05)
  filter(measure %in% c("cbf","gm_md","qt1")) |>
  mutate(acro=ifelse(qvalue<0.05,acrophase,NA)) |> 
  select(measure,roi,acro)
write_csv(fig3bSD |> mutate(measure=int2extlab[measure]),
          "output/source_data/Figure_3b_part1.csv")

# GM-MD and GM-qT1 are reordered in post
spatial_p <-
  inner_join(fig3bSD,meta) |> 
  ungroup() |>
  ggplot(aes(fill=acro))+
  facet_wrap(~int2extlab[measure],nrow=5)+
  theme_condense()+
  scale_fill_clock(rot=15,na.value="grey95")+
  geom_brain(atlas = glasser,
             position=position_brain(side ~ hemi),
             color="grey",size=0.1)+
  theme_void() 
spatial_p
```

## Subcortical

```{r}
myaseg <- get_subcor_ggseg_atlas()
spatial_subcort_p <- fig3bSD |>  
	rename(label=roi) |> 
	ungroup() |> 
	select(label,acro,measure) |> 
	ggplot(aes(fill=acro))+
	theme_condense()+
	scale_fill_clock(rot=15,na.value="grey95")+
	facet_wrap(~int2extlab[measure],nrow=3)+
	geom_brain(atlas = myaseg)+
	theme_void()
spatial_subcort_p
```

## ICBM (WM-MD)

To be added manually to final figure

```{r}
library(ggsegICBM)
library(plotly)
library(ggseg3d)
meta <- get_icbm_spatial_meta()
pdf <- ROI_Gcosinor |> filter(measure=="md") 

# Source Data
fig3bSD2 <- 
  pdf |> 
  mutate(acro=ifelse(qvalue<0.05,acrophase,NA)) |> 
  select(measure,roi,acro)
write_csv(fig3bSD2 |> mutate(measure=int2extlab[measure]),
          "output/source_data/Figure_3b_part2.csv")

pdf2 <- inner_join(fig3bSD2,meta) |> icbm_preprocess()
pdf2$region %in% icbm_3d$ggseg_3d[[1]]$region

rot=15
pal <- 1:24
names(pal) <- rainbow(24)[((1:24)-1+rot)%%24+1]


p3d <- 	
  pdf2 |> 
  select(measure,acro,roi,region) |> 
	ggseg3d(atlas = icbm_3d,colour="acro",
			palette = pal,na.alpha = 0.01) |> 
	add_glassbrain(opacity=0.05) |>
	pan_camera("right lateral") |>
	remove_axes()
p3d

pdf2 |> 	
	select(region,acro,roi) |> 
	ggseg3d(atlas = icbm_3d,colour="acro",
			palette = pal,na.alpha = 0.01) |> 
	add_glassbrain(opacity=0.05) |>
	pan_camera("left lateral") |>
	remove_axes()
```

## Color legend

```{r}
circcols_p <-
	predat |> 
	ggplot()+
	geom_rect(xmin=-1,xmax=25,ymin=-1,ymax=0,fill="white",color=NA)+
	coord_polar()+
	scale_y_continuous(limits = c(-1,NA))+
	scale_x_continuous(limits=c(0,24),breaks=seq(0,24,3))+
	xlab("ROI Population-Mean Acrophase")+
	ylab("")+
	theme_minimal()+theme_condense()+
	theme(
		panel.grid.minor.x = element_blank(),
		  panel.grid.minor.y = element_blank(),
		  panel.grid.major.y = element_blank(),
		  panel.grid.major.x = element_blank()
		)+
	theme(axis.text.x = element_text(size=6),
		  axis.text.y = element_blank())+
	geom_tile(data=data.frame(ymax=rep(2,200),
							  x=c(seq(0,24,length.out=200))),
			  aes(x,y=ymax, fill = x),
			  color=NA,alpha=0.9) +
	scale_fill_clock(rot=15,guide="none")
circcols_p
```

# Figure 3 {.tabset}

## Figure 3

```{r figure_3, fig.height=12,fig.width=10, dev=c("png","pdf")}
theme_condense <- function(){
	list(
		theme(strip.background = element_rect(color='white',fill='white')),
		theme(strip.text = element_text(size=6,color="black",
										margin = margin( b = 0, t = 0) ) ),
		theme(axis.text = element_text(size=6))
		)
}

library(patchwork)
library(ggpubr)
f3a_sub_act_ss +ggtitle("a")+theme(strip.text = element_blank())+
	#
	rhosum_ss_p+theme_minimal()+theme_condense()+theme(axis.text.x = element_text())+theme(strip.text = element_blank())+
	geom_hline(yintercept = c(0,1),color="grey80")+
	theme(axis.text.y = element_blank())+
	#
	spatial_p + theme(legend.position = "none")+ggtitle("b")+theme(strip.text = element_blank())+
	#
	gcrose_ss_p + theme(legend.position = 'none')+theme_condense()+ ggtitle("c")+
	theme(axis.text.y = element_text(size=6))+theme(strip.text = element_text(size=14,color="black"))+
	#
	as_ggplot(get_legend(gcrose_ss_p))+
	circcols_p+
	# annotate("text",0,-0.7,label="Population-Mean Acrophase")+
	#patchwork
	plot_layout(design = "
				AAABCCCCCDDDD
				AAABCCCCCDDDD
				AAABCCCCCDDDD
				AAABEEEEEDDDD
				####FFFFF####
				")+
	# plot_annotation(tag_levels = 'a')+
	plot_layout(guides = 'collect')
```

## Figure S2 (Rest of 3)

```{r figure_rest_of_3, dev=c("png","pdf"), fig.width=8,fig.height=10}
library(patchwork)
f3a_sub_act + theme_condense()+
  rhosum_p +theme_condense()+
  theme(axis.text.y = element_text(size=6))+
  p_roiact +  theme_condense()+
	gcrose_p + theme_condense()+
	#patchwork
	plot_layout(design = "AABBCCDD")+
	plot_annotation(tag_levels = 'a')+
	plot_layout(guides = 'collect')
```


